Eye-tracking on the web: lessons learned from replicating 5 experiments

### 2022-04-11 manuscript.tex
update E3 analysis

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  man,floatsintext]{apa6}
\title{Eye-tracking on the web: lessons learned from replicating 5 experiments}
\author{Josh de Leeuw\textsuperscript{1} \& et al.\textsuperscript{1,2}}
\date{}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Eye-tracking on the web: lessons learned from replicating 5 experiments},
  pdfauthor={Josh de Leeuw1 \& et al.1,2},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Just picked a random title: feel free to change}
\keywords{keywords\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifXeTeX
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

@@

\abstract{
ADD LATER
}

@@

We used R {[}Version 4.1.2; R Core Team (2021){]} and the R-packages \emph{afex} {[}Version 1.0.1; Singmann, Bolker, Westfall, Aust, and Ben-Shachar (2021){]}, \emph{broom.mixed} {[}Version 0.2.7; Bolker and Robinson (2020){]}, \emph{dplyr} {[}Version 1.0.7; Wickham, François, Henry, and Müller (2021){]}, \emph{forcats} {[}Version 0.5.1; Wickham (2021a){]}, \emph{ggplot2} {[}Version 3.3.5; Wickham (2016){]}, \emph{jsonlite} {[}Version 1.7.2; Ooms (2014){]}, \emph{lme4} {[}Version 1.1.27.1; Bates, Mächler, Bolker, and Walker (2015){]}, \emph{lmerTest} {[}Version 3.1.3; Kuznetsova, Brockhoff, and Christensen (2017){]}, \emph{Matrix} {[}Version 1.3.4; Bates and Maechler (2021){]}, \emph{papaja} {[}Version 0.1.0.9997; Aust and Barth (2020){]}, \emph{readr} {[}Version 2.1.1; Wickham and Hester (2020){]}, \emph{shiny} {[}Version 1.7.1; Chang et al. (2021){]}, \emph{stringr} {[}Version 1.4.0; Wickham (2019){]}, \emph{tidyr} {[}Version 1.1.4; Wickham (2021b){]}, and \emph{tinylabels} (Barth, 2022) for all our analyses.

@@

@@

\textbf{Eye-gaze}. Looks during the retrieval period were categorized as belonging to one of four quadrants based on the x,y coordinates. The critical quadrant was the one in which the to-be-retrieved object had been previously located during encoding. The other three quadrants were semi-randomly labeled ``first,'' ``second,'' third'' (e.g., when the critical quadrant was in the top left, the ``first'' quadrant was the top right quadrant, but when the critical quadrant was in the top right, ``first'' corresponded to bottom right, etc.). In both the fixed- and free-viewing condition, participants directed a larger proportion of looks to the critical quadrant (see Figure \ref{fig:E2-gaze-fig-both-conds}). This bias appeared larger in the free-viewing condition, suggesting that the manipulation was (somewhat) effective.

@@

@@

@@

@@

@@

\leavevmode\vadjust pre{\hypertarget{ref-papoutsaki2016webgazer}{}}%
Papoutsaki, A., Sangkloy, P., Laskey, J., Daskalova, N., Huang, J., \& Hays, J. (2016). {WebGazer}: {Scalable} webcam eye tracking using user interactions. In \emph{Proceedings of the 25th international joint conference on artificial intelligence ({IJCAI})} (pp. 3839--3845). {AAAI}.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Eye-tracking on the web: lessons learned from replicating 5 experiments},
  pdfauthor={Josh de Leeuw1 \& et al.1,2},
  pdflang={en-EN},
  pdfkeywords={keywords},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

@@

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{keywords\newline\indent Word count: X}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Eye-tracking on the web: lessons learned from replicating 5 experiments}
\author{Josh de Leeuw\textsuperscript{1} \& et al.\textsuperscript{1,2}}
\date{}

\shorttitle{Just picked a random title: feel free to change}

@@

\abstract{%
ADD LATER
}

@@

We used R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{afex} (Version 1.1.0; Singmann, Bolker, Westfall, Aust, \& Ben-Shachar, 2021), \emph{broom.mixed} (Version 0.2.9.3; Bolker \& Robinson, 2020), \emph{dplyr} (Version 1.0.8; Wickham, François, Henry, \& Müller, 2021), \emph{forcats} (Version 0.5.1; Wickham, 2021a), \emph{ggplot2} (Version 3.3.5; Wickham, 2016), \emph{jsonlite} (Version 1.8.0; Ooms, 2014), \emph{lme4} (Version 1.1.28; Bates, Mächler, Bolker, \& Walker, 2015), \emph{lmerTest} (Version 3.1.3; Kuznetsova, Brockhoff, \& Christensen, 2017), \emph{Matrix} (Version 1.3.4; Bates \& Maechler, 2021), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{readr} (Version 2.1.2; Wickham \& Hester, 2020), \emph{shiny} (Chang et al., 2021), \emph{stringr} (Version 1.4.0; Wickham, 2019), \emph{tidyr} (Version 1.2.0; Wickham, 2021b), and \emph{tinylabels} (Version 0.2.3; Barth, 2022) for all our analyses.

@@

\begin{verbatim}
## Warning: package 'patchwork' was built under R version 4.1.3
\end{verbatim}

@@

\textbf{Eye-gaze}. Looks during the retrieval period were categorized as belonging to one of four quadrants based on the x,y coordinates. The critical quadrant was the one in which the to-be-retrieved object had been previously located during encoding. The other three quadrants were semi-randomly labeled ``first'', ``second,'' third'' (e.g., when the critical quadrant was in the top left, the ``first'' quadrant was the top right quadrant, but when the critical quadrant was in the top right, ``first'' corresponded to bottom right, etc.). In both the fixed- and free-viewing condition, participants directed a larger proportion of looks to the critical quadrant (see Figure \ref{fig:E2-gaze-fig-both-conds}). This bias appeared larger in the free-viewing condition, suggesting that the manipulation was (somewhat) effective.

@@

\hypertarget{day-1}{%
\paragraph{Day 1}\label{day-1}}

During day 1 of the experiment, participants viewed pairs of images, one of which was always familiar and the other unfamiliar. We calculated a looking score for each participant, defined as the proportion of gaze samples in the ROI of the unfamiliar image out of all the gaze samples that were in either ROI. Gaze samples that were not in either ROI were not included in this analysis. A looking score of 0.5 indicates that participants looked equally often at the familiar and unfamiliar images, while a looking score above 0.5 indicates a preference for the unfamiliar object and a looking score below 0.5 indicate a preference for the familiar object.

Of the 1248 trials in the experiment, 78 had no fixations in either ROI, and so the looking score was unknown. We removed these trials from this analysis.

The mean looking score was 0.55 (\emph{SD} = 0.10). This significantly greater than 0.5, \emph{t}(49) = 3.29, \emph{p} = 0.00, indicating that participants did show a preference for looking at the novel objects.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Plot of cumulative looking score-1.pdf}
\caption{(\#fig:Plot of cumulative looking score)Cumulative looking score over the 5 second exposure during part 2 of day 1. Error bars represent +/- 1 SEM.}
\end{figure}

\hypertarget{day-2}{%
\paragraph{Day 2}\label{day-2}}

In all of these analyses, we excluded the 16 (out of 2304) trials where the response time for the recognition judgment was greater than 10 seconds.

Participants correctly identified whether the image was familiar or unfamiliar 87.09\% (\emph{SD} = 10.49) of the time. After excluding the 3 participants who responded correctly to all images, the average confidence rating for correct responses (M = 3.51; SD = 0.41) was significantly higher than their average confidence ratings for incorrect responses (M = 2.55; SD = 0.75), t(44) = -9.36, p = 0.00 . Among the same subset of participants, response times for correct responses (M = 1,443.49, SD = 413.94) were also significantly faster than for incorrect responses (M = 2,212.65, SD = 1,733.76), t(44) = 3.43 , p = 0.00.

To see whether preferentially looking an the unfamiliar object on day 1 was correlated with confidence and response time for correct responses on day 2, we computed the correlation coefficient between day 1 looking scores and day 2 confidence/RT for each participant. Following the original analysis, we transformed these values using the Fisher p-to-z transformation. Using one-sample t-tests, we found no significant different from 0 for the correlation between looking score and confidence ratings, t(38) = 0.46, p = 0.65 (excluding the subjects who gave the same confidence judgment for all images), nor the the correlation between looking score and RT, t(46) = 0.49, p = 0.63.

\begin{verbatim}
## Warning: Removed 75 rows containing missing values (geom_point).
\end{verbatim}

\includegraphics{manuscript_files/figure-latex/Plot Looking Score Correlations-1.pdf}

In the original experiment, the two objects on day 1 were presented on two separate monitors and gaze was coded by manually coding video recordings. In our replication analysis, we analyzed eye movement data using ROIs defined around the two images. In this section we explore an alternative coding of the eye movement data by coding simply left half vs.~right half of the screen. The coarser coding may be more appropriate for webcam-based eyetracking.

The correlation between looking scores using the ROI method and the halves method is 0.76.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/E3-roi correlation of looking score-1.pdf}
\caption{(\#fig:E3-roi correlation of looking score)Correlation between looking scores calculated using ROIs and using screen halves.}
\end{figure}

\hypertarget{looking-scores}{%
\paragraph{Looking Scores}\label{looking-scores}}

When looking scores are coded as left vs.~right half of the screen, we find that participants looked more at the novel object. The mean looking score was 0.54 (\emph{SD} = 0.08). This was significantly greater than 0.5, \emph{t}(50) = 3.51, \emph{p} = 0.00.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/E3-roi Plot of cumulative looking score-1.pdf}
\caption{(\#fig:E3-roi Plot of cumulative looking score)Cumulative looking score over the 5 second exposure during part 2 of day 1. Error bars represent +/- 1 SEM.}
\end{figure}

\hypertarget{correlations-with-day-2-performance}{%
\paragraph{Correlations with Day 2 Performance}\label{correlations-with-day-2-performance}}

Performance on day 2 remained uncorrelated with day 1 looking scores after switching the coding of gaze. We found no significant different from 0 for the correlation between looking score and confidence ratings, t(39) = 0.74, p = 0.47 (excluding the subjects who gave the same confidence judgment for all images), nor the the correlation between looking score and RT, t(47) = 0.28, p = 0.78.

To see if calibration success is correlated with the eye tracking effects, we calculated a calibration score for each participant. The calibration score was the average proportion of samples within XXX pixels of the validation points during the final validation phase before the eye tracking is performed.

Calibration scores were not correlated with looking scores, regardless of which method was used to calculate looking scores.

\includegraphics{manuscript_files/figure-latex/E3-cal Plot looking score (halves) by calibration-1.pdf}

\includegraphics{manuscript_files/figure-latex/E3-cal Plot looking score (roi) by calibration-1.pdf}

We then looked at the correlation of calibration scores with the correlation between day 2 memory performance and day 1 looking scores for both kinds of behavioral and looking measures. None of the four relationships showed a significant correlation.

\includegraphics{manuscript_files/figure-latex/E3-cal Plot correlations with calibration.score-1.pdf}

@@

\begin{verbatim}
## Warning: Computation failed in `stat_summary()`:
\end{verbatim}

@@

\hypertarget{calibration-3}{%
\subsubsection{Calibration}\label{calibration-3}}

@@

\hypertarget{effects-of-rois-1}{%
\subsubsection{Effects of ROIs}\label{effects-of-rois-1}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-papoutsaki2016webgazer}{}}%
Papoutsaki, A., Sangkloy, P., Laskey, J., Daskalova, N., Huang, J., \& Hays, J. (2016). {WebGazer}: {Scalable} webcam eye tracking using user interactions. \emph{Proceedings of the 25th International Joint Conference on Artificial Intelligence ({IJCAI})}, 3839--3845. {AAAI}.
